\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{mathtools}
\usepackage{geometry}
\geometry{margin=2.5cm}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{amsfonts}
\usepackage{amsopn}
\usepackage{doi}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{hypothesis}[theorem]{Hypothesis}

\newcommand{\Rn}{R_n}
\newcommand{\Sn}{\bar{\Sigma}_n}
\newcommand{\dexact}{\hat{\delta}_{\mathrm{exact}}}
\newcommand{\deff}{\delta_{\mathrm{eff}}}
\newcommand{\Reff}{R_n^{\mathrm{eff}}}
\newcommand{\bstar}{\beta^{*}}

\title{Observable-Class Derivation Mismatch and the Exact Growth Equation:\\
From Proxy Capacity to Block-Level Dynamics}
\author{J\'er\^ome Beau\\
\small Independent Researcher, France; jerome.beau@cosmochrony.org}
\date{2026}

\begin{document}
\maketitle

\begin{abstract}
The O3--O7 derivation chain establishes the structural relation
$\bstar = 1/(\delta + 1/2)$ linking the capacity decay exponent~$\delta$
to the cascade growth exponent~$\bstar$ governing charged-lepton mass ratios.
Papers O12--O14 demonstrated that the exact Weil-block capacity exponent
$\dexact(q) \in [3.6, 4.8]$ is incompatible with the target window
$\delta \in [7.4, 10.6]$, and that this tension is structural rather than
a finite-size artefact.
The present paper identifies the precise origin of the mismatch:
the O6 growth equation $\mathrm{d}p/\mathrm{d}n \lesssim \Rn \, c_{\mathrm{BI}} \sqrt{p(n)}$
derives $\bstar$ from a \emph{scalar global} redundancy functional~$\Rn$,
whereas the exact observable measured in O12--O13 is a \emph{block-averaged}
capacity over $(q-1)$ Weil representations of internal dimension~$q$.
We prove a non-equivalence theorem establishing that the block-averaged
exponent $\dexact$ does not, in general, coincide with the scalar exponent~$\alpha$
entering the growth equation.
We define the effective dynamical redundancy $\Reff$ as the observable
that genuinely governs the valence growth in the exact-block setting,
and re-derive the growth equation with this observable as input.
The re-derivation yields two structural scenarios:
(C1)~the functional form $\bstar = 1/(\deff + 1/2)$ survives with a
corrected exponent $\deff$ that differs from $\dexact$ by an
aggregation bias;
(C2)~the functional form itself is modified by a
representation-dimension correction,
$\bstar = 1/(\deff + 1/2 + \sigma(q))$, where $\sigma(q) \to 0$
as $q \to \infty$.
Scenario discrimination requires a block-level redundancy measurement
not yet available from the O12 pipeline,
identified as the central open direction of this paper.
\end{abstract}

\smallskip\noindent
\textbf{Keywords.}
Observable-class mismatch, growth equation, block-averaged capacity,
scalar redundancy, Weil representations, Heisenberg graphs,
cascade exponent, spectral admissibility.

  \clearpage
\tableofcontents

\clearpage
\subsection*{Notation glossary}

  \noindent
  The four exponents appearing in this paper form a strict hierarchy:
  \begin{center}
    \begin{tabular}{clp{7.5cm}}
      \toprule
      Symbol & Name & Definition \\
      \midrule
      $\dexact$ & measured exponent &
      Power-law fit of the unweighted block average
      $\Sn = \mathbb{E}_c[\Sigma_n^{(c)}] \sim A\,n^{-\dexact}$
      (O12/O13 observable). \\[4pt]
      $\alpha_{\mathrm{dyn}}$ & dynamical exponent &
      Power-law fit of the weighted observable
      $\Reff \sim B\,n^{-\alpha_{\mathrm{dyn}}}$
      (Definition~\ref{def:alphaeff}).
      Satisfies $\alpha_{\mathrm{dyn}} \leq \dexact$
      (Corollary~\ref{cor:nogo}). \\[4pt]
      $\deff$ & effective exponent (C1) &
      Redefined exponent absorbing shell-growth and
      $q$-normalisation corrections:
      $\deff = \dexact - (D-1) + \log q / \log n^{*}$. \\[4pt]
      $\sigma(q)$ & structural correction (C2) &
      Additional $q$-dependent term in the denominator
      of the growth formula, required if C1 does not
      close the gap to $\bstar \in (0.09, 0.13)$. \\
      \bottomrule
    \end{tabular}
  \end{center}

\section{Introduction}
\label{sec:intro}

\subsection{Position in the O-series}
\label{sec:position}

The admissibility sub-programme has traced a complete arc from obstruction
identification (O8--O9~\cite{Beau2026a12,Beau2026a13}) through exact
measurement (O10--O13~\cite{Beau2026a14,Beau2026a15,Beau2026a16,Beau2026a17})
to theoretical correspondence (O14~\cite{Beau2026a18}).
O14 established that the tension between the exact Weil-block capacity
exponent $\dexact(q) \in [3.6, 4.8]$ and the phenomenological target
$\delta \in [7.4, 10.6]$ is not a numerical artefact (refuting hypothesis~S1
of O13) but a structural mismatch (Source~S2).
The corrected relation derived in O14,
\begin{equation}
\label{eq:O14corrected}
\bstar = \frac{1}{\deff + \tfrac{1}{2}} + \epsilon(q),
\end{equation}
with $\deff = \dexact - \delta_{\gamma,\mathrm{corr}}
- \eta \cdot \log q / \log n^{*}$ and $\epsilon(q) \to 0$ as $q \to \infty$,
was shown to be consistent with Scenario~S2-B for all tested primes:
$\deff < 5.0$ throughout, with the normalisation correction
$\eta \cdot \log q / \log n_1 \approx 1.03$--$1.06$ as the dominant term.

O14 thus closed the measurement question but left open the
\emph{derivation} question:
does the functional form $\bstar = 1/(\delta + 1/2)$ itself remain valid
when the input observable changes from a scalar global capacity to a
block-averaged Weil capacity?
The present paper answers this question.

\subsection{The central question}
\label{sec:central}

The relation $\bstar = 1/(\delta + 1/2)$ was derived in
O6~\cite{Beau2026a10} (Proposition~4.2) from the growth equation
\begin{equation}
\label{eq:growth}
\frac{\mathrm{d}p}{\mathrm{d}n}
\lesssim \Rn \cdot c_{\mathrm{BI}} \sqrt{p(n)},
\end{equation}
where $\Rn$ is a scalar redundancy functional satisfying
$\Rn \sim p(n)^{-\alpha}$ in the pre-saturation window.
Separation of variables yields $p(n) \sim n^{1/(\alpha + 1/2)}$,
giving $\bstar = 1/(\alpha + 1/2)$.

The question is: \emph{what is the correct identification of~$\Rn$
in the exact Weil-block setting, and does substituting it into~\eqref{eq:growth}
preserve the functional form?}

\subsection{Main results}
\label{sec:mainresults}

\begin{enumerate}[label=(\arabic*)]
\item We audit the O3--O7 derivation chain and identify the three
precise points where the scalar-observable hypothesis enters
(Section~\ref{sec:audit}).
\item We prove a non-equivalence theorem: the block-averaged exponent
$\dexact$ does not coincide with the scalar exponent~$\alpha$ of the
growth equation under generic conditions on block heterogeneity
(Section~\ref{sec:noneq}).
\item We define the effective dynamical redundancy $\Reff$ and show it
is the correct input to~\eqref{eq:growth} in the exact setting
(Section~\ref{sec:Reff}).
\item We re-derive the growth equation with $\Reff$ as input and identify
the two structural scenarios C1 and C2 (Section~\ref{sec:rederivation}).
\item We specify the block-level measurement required to discriminate
C1 from C2 (Section~\ref{sec:discrimination}).
\end{enumerate}

\section{Audit of the O3--O7 Derivation Chain}
\label{sec:audit}

We trace the logical chain that produces $\bstar = 1/(\delta + 1/2)$
and identify, at each step, the hypotheses that hold for the proxy
observable but fail for the exact Weil-block observable.

\subsection{Step A: O3 mass formula}
\label{sec:stepA}

O3~\cite{Beau2026a7} derives the inter-generation mass ratio
\begin{equation}
\label{eq:massratio}
\frac{M_i}{M_j}
= \left(\frac{|{\lambda_i - 1}|}{|{\lambda_j - 1}|}\right)^{2/\beta},
\end{equation}
from the support-narrowing mechanism under the power-law growth
hypothesis $p(n) \sim A n^\beta$.
The derivation uses two hypotheses:
\begin{description}[style=nextline]
\item[H-O3-1] The growth $p(n) \sim A n^\beta$ is a power law.
\item[H-O3-2] The mass is inversely proportional to the exit rank:
$M_i \propto E_P / n_{\mathrm{exit}}(\lambda_i)$.
\end{description}
Both hypotheses are explicit structural postulates of O3.
Neither involves any property of the capacity observable.
The mass formula~\eqref{eq:massratio} is therefore \emph{independent}
of the observable-class mismatch and remains valid throughout.

\subsection{Step B: O6 growth equation}
\label{sec:stepB}

O6~\cite{Beau2026a10}, Proposition~4.2, derives $\bstar = 1/(\alpha + 1/2)$
from the growth equation~\eqref{eq:growth}.
The derivation involves three implicit hypotheses about the
redundancy functional~$\Rn$:

\begin{description}[style=nextline]
\item[H-O6-1 (Scalar observable)]
$\Rn$ is a single real-valued function of the cascade rank~$n$,
independent of any internal representation label.
\item[H-O6-2 (Power-law decay)]
$\Rn \sim p(n)^{-\alpha}$ for a single exponent~$\alpha$ over the
pre-saturation window.
\item[H-O6-3 (Direct coupling)]
$\Rn$ enters the growth equation multiplicatively:
the rate of valence increase is proportional to~$\Rn$.
\end{description}

Hypothesis H-O6-1 is the central point of failure.
In the proxy setting (O7--O11), the capacity $C_n^{\mathrm{proxy}}$
is defined via a Fourier-character proxy that abelianises the
representation structure~\cite{Beau2026a11}.
It is indeed a scalar function of~$n$.

In the exact setting (O12--O14), the observable is
$\Sn = \mathbb{E}_c[\Sigma_n^{(c)}]$, a mean over $(q-1)$ Weil blocks,
each of internal dimension~$q$.
This is not a scalar observable in the sense of H-O6-1: it carries
a hidden dependence on the block structure through the averaging.

\subsection{Step C: O7 reinterpretation}
\label{sec:stepC}

O7~\cite{Beau2026a11} recasts the O6 formula in capacity language.
The identification $\alpha = \delta$ (O7, Section~5.1) holds in the
reduced filling model where $p(n) = n$.
The state law $R_n^{(k)} \approx \Phi(\eta_n)$ with
$\Phi(\eta) = 1/\sqrt{1 + \eta^2}$ is derived exactly in the reduced
model and is a property of the scalar capacity.

O7 does not introduce new deductive content for the $\delta \to \bstar$
relation.
It inherits H-O6-1 from O6 without modification.
The Remark~5.2 of O7 states this explicitly:
the formula ``does not introduce a new derivation of~$\beta$:
it recasts the O6 target in capacity language without circularity.''

\subsection{Summary of audit}
\label{sec:auditsummary}

\begin{table}[ht]
\centering
\caption{Audit of the derivation chain.
The column ``Exact status'' indicates whether the hypothesis
holds for the exact Weil-block observable.}
\label{tab:audit}
\begin{tabular}{llcc}
\toprule
Paper & Hypothesis & Proxy & Exact \\
\midrule
O3 & H-O3-1 (power-law growth) & valid & valid \\
O3 & H-O3-2 (mass $\propto 1/n_{\mathrm{exit}}$) & valid & valid \\
O6 & H-O6-1 (scalar observable) & valid & \textbf{fails} \\
O6 & H-O6-2 (power-law decay) & valid & valid \\
O6 & H-O6-3 (direct coupling) & valid & open \\
O7 & inherits H-O6-1 & valid & \textbf{fails} \\
\bottomrule
\end{tabular}
\end{table}

The mass formula of O3 is intact.
The growth equation of O6 is algebraically correct in its domain.
The failure is localised at H-O6-1: the scalar-observable hypothesis,
which enters the growth equation through the identification of~$\Rn$
with a single power-law observable.

\section{Non-Equivalence of Block-Averaged and Scalar Exponents}
\label{sec:noneq}

We now prove that the exponent extracted from a block-averaged observable
does not, in general, coincide with the exponent of a scalar observable
satisfying the growth equation.

\subsection{Setup}

Fix an odd prime~$q$.
The Heisenberg group $G_q = \mathrm{Heis}_3(\mathbb{Z}/q\mathbb{Z})$
has $(q-1)$ faithful irreducible representations
$\rho_c$ ($c \in (\mathbb{Z}/q\mathbb{Z})^\times$), each of dimension~$q$.
The block-level projective capacity at BFS depth~$n$ in block~$c$ is
\begin{equation}
\label{eq:blockcap}
\Sigma_n^{(c)} = \frac{\Delta r_n^{(c)}}{|S_n|},
\end{equation}
where $\Delta r_n^{(c)}$ is the incremental projective span in
representation~$\rho_c$ and $|S_n|$ is the BFS shell size.
The block average is $\Sn = \mathbb{E}_c[\Sigma_n^{(c)}]$.

In the proxy setting, the capacity is a single scalar
$C_n^{\mathrm{proxy}}$ defined via the Fourier-character
abelianisation~\cite{Beau2026a15}.

\subsection{Block heterogeneity and exponent bias}

\begin{definition}[Block exponent and mean exponent]
\label{def:exponents}
For each block~$c$, define the block exponent
$\delta^{(c)}$ by the power-law fit
$\Sigma_n^{(c)} \sim A_c \, n^{-\delta^{(c)}}$
over the fitting window $[n_0, n_1]$.
Define the mean exponent
$\bar{\delta} = \mathbb{E}_c[\delta^{(c)}]$
and the exponent of the mean
$\dexact$ by the fit
$\Sn \sim A \, n^{-\dexact}$.
\end{definition}

\begin{proposition}[Non-equivalence]
\label{prop:noneq}
If the block amplitudes $A_c$ are not all equal,
then $\dexact \neq \bar{\delta}$ in general.
Specifically, let $V_A = \mathrm{Var}_c(A_c) / \mathbb{E}_c[A_c]^2$
be the relative amplitude variance.
Then
\begin{equation}
\label{eq:noneq}
\dexact = \bar{\delta}
- \frac{1}{\log(n_1/n_0)}
\log\!\left(
\frac{\mathbb{E}_c[A_c \, (n_1/n_0)^{-\delta^{(c)}}]}
{\mathbb{E}_c[A_c] \cdot (n_1/n_0)^{-\bar{\delta}}}
\right).
\end{equation}
The correction term vanishes if and only if $\delta^{(c)}$ is independent
of~$c$ (block universality).
\end{proposition}

\begin{proof}
Write $\Sn = \mathbb{E}_c[A_c \, n^{-\delta^{(c)}}]$.
At two endpoints $n_0, n_1$ of the fitting window:
\[
\dexact
= -\frac{\log(\Sn[n_1] / \Sn[n_0])}{\log(n_1 / n_0)}
= -\frac{1}{\log(n_1/n_0)}
\log\!\left(
\frac{\mathbb{E}_c[A_c \, n_1^{-\delta^{(c)}}]}
{\mathbb{E}_c[A_c \, n_0^{-\delta^{(c)}}]}
\right).
\]
If all $\delta^{(c)} = \delta_0$, this reduces to $\delta_0 = \bar{\delta}$.
Otherwise, the ratio is a weighted average of $\exp(-\delta^{(c)} \log(n_1/n_0))$
with weights $A_c$, and by Jensen's inequality it differs from
$\exp(-\bar{\delta} \log(n_1/n_0))$ whenever the distribution of
$\delta^{(c)}$ is non-degenerate.
\end{proof}

\begin{remark}[Direction of the bias]
\label{rem:bias}
For a convex function $f(\delta) = \exp(-\delta \log(n_1/n_0))$,
Jensen's inequality gives
$\mathbb{E}_c[f(\delta^{(c)})] \geq f(\bar{\delta})$,
so $\dexact \leq \bar{\delta}$.
The block-averaged exponent is systematically \emph{smaller}
than the mean of individual block exponents.
This is consistent with the O12--O13 observation that $\dexact$
lies below the target window.
\end{remark}

\begin{remark}[Quantitative estimate from O12--O13 data]
\label{rem:quantitative}
O12~\cite{Beau2026a16} reports inter-block variance ratios
$V_n^{\max}(q)$ that decrease from $\sim 0.15$ at $q = 29$ to
$\sim 0.03$ at $q = 211$.
Even at $q = 211$, the residual variance produces a bias of order
$|\dexact - \bar{\delta}| \sim 0.1$--$0.3$, which is insufficient
to bridge the gap to the target $\delta \in [7.4, 10.6]$.
The non-equivalence is therefore real but quantitatively small
compared to the total deficit.
\end{remark}

\subsection{Consequence for the growth equation}

The growth equation~\eqref{eq:growth} couples the rate of valence
increase to the redundancy of new relational directions.
Physically, it is the \emph{fraction of genuinely new directions
contributed by the BFS frontier} that limits the growth of~$p$.
This fraction is a property of the \emph{aggregate} projective
dynamics, not of any individual block.

The scalar exponent~$\alpha$ in H-O6-2 is the exponent of this
aggregate fraction.
It is not, in general, equal to $\dexact$ (the exponent of the
block-averaged capacity), because the block-averaged capacity
includes internal normalisation and phase structure that do not
enter the growth equation.

This motivates the definition of the effective dynamical redundancy
in the next section.

\section{Effective Dynamical Redundancy}
\label{sec:Reff}

\subsection{Physical meaning of the growth equation}

The growth equation~\eqref{eq:growth} has a precise physical content:
at each cascade step~$n$, the substrate attempts to increase its
effective relational valence~$p$ by incorporating new relational
directions from the BFS frontier.
The rate of successful incorporation is limited by two factors:
\begin{enumerate}[label=(\alph*)]
\item The Born--Infeld flux bound $c_{\mathrm{BI}}$, which caps the rate
of change of any relational variable~\cite{Beau2026a8}.
\item The fraction of frontier directions that are genuinely
new, i.e.\ not already represented in the current relational
structure.
\end{enumerate}
Factor~(b) is what $\Rn$ encodes.
In the proxy setting, where the capacity is a scalar, $\Rn$ is
directly proportional to the capacity: $\Rn \propto \Sigma_n$.
In the exact setting, the identification is more subtle because
``genuinely new'' must be defined relative to the full block structure.

\subsection{Definition of $\Reff$}

\begin{definition}[Effective dynamical redundancy]
\label{def:Reff}
The effective dynamical redundancy at cascade rank~$n$ is
\begin{equation}
\label{eq:Reffdef}
\Reff = \frac{1}{q-1} \sum_{c \in (\mathbb{Z}/q\mathbb{Z})^\times}
w_c(n) \, \Sigma_n^{(c)},
\end{equation}
where $w_c(n)$ is the dynamical weight of block~$c$, defined as
\begin{equation}
\label{eq:weights}
w_c(n)
= \frac{\Delta r_n^{(c)} / \dim V_c}
{\sum_{c'} \Delta r_n^{(c')} / \dim V_{c'}}
= \frac{\Delta r_n^{(c)} / q}
{\sum_{c'} \Delta r_n^{(c')} / q}
= \frac{\Delta r_n^{(c)}}{\sum_{c'} \Delta r_n^{(c')}}.
\end{equation}
\end{definition}

The weight $w_c(n)$ measures the relative contribution of block~$c$
to the total incremental span at step~$n$.
It is the natural dynamical weight because the growth equation
counts the total number of new relational degrees of freedom,
and each block contributes $\Delta r_n^{(c)}$ such degrees.

\begin{proposition}[Relation to the unweighted mean]
\label{prop:relation}
Let $\Sn = \mathbb{E}_c[\Sigma_n^{(c)}]$ be the unweighted block
average (the O12/O13 observable) and let
$\Sigma_n^{w} = \sum_c w_c \Sigma_n^{(c)}$ be the weighted observable
of Definition~\ref{def:Reff}.
Then
\begin{equation}
\label{eq:relation}
\Reff = \Sn + \mathrm{Cov}_c\!\left(
\frac{w_c}{1/(q-1)}, \, \Sigma_n^{(c)}
\right),
\end{equation}
where the covariance is taken over the uniform distribution on blocks.
The correction vanishes if the weights are uniform ($w_c = 1/(q-1)$
for all~$c$), which occurs if and only if all blocks contribute
equally to the incremental span.
\end{proposition}

\begin{proof}
Direct computation:
$\Reff - \Sn
= \sum_c w_c \Sigma_n^{(c)} - \frac{1}{q-1} \sum_c \Sigma_n^{(c)}
= \sum_c (w_c - \frac{1}{q-1}) \Sigma_n^{(c)}
= (q-1) \mathrm{Cov}_c(w_c, \Sigma_n^{(c)})$.
\end{proof}

\begin{remark}[Sign of the correction]
\label{rem:sign}
Blocks with larger $\Delta r_n^{(c)}$ (hence larger $w_c$) tend to
have larger $\Sigma_n^{(c)}$ (since $\Sigma_n^{(c)}$ is the ratio
$\Delta r_n^{(c)} / |S_n|$).
The covariance is therefore generically positive, giving
$\Reff \geq \Sn$.
The effective dynamical redundancy is \emph{larger} than the
unweighted block average.
\end{remark}

\subsection{Exponent of $\Reff$}

\begin{definition}[Dynamical exponent]
\label{def:alphaeff}
The dynamical exponent~$\alpha_{\mathrm{dyn}}$ is defined by the
power-law fit $\Reff \sim B \, n^{-\alpha_{\mathrm{dyn}}}$ over the
pre-saturation window.
\end{definition}

By Proposition~\ref{prop:relation} and Remark~\ref{rem:sign},
$\alpha_{\mathrm{dyn}} \leq \dexact$: the dynamical exponent is
smaller than or equal to the unweighted block-average exponent.

\begin{corollary}[Aggregation no-go]
\label{cor:nogo}
No choice of dynamical weights $w_c(n)$ satisfying $w_c \geq 0$
and $\sum_c w_c = 1$ can produce
$\alpha_{\mathrm{dyn}} > \dexact$.
In particular, the block-averaged exponent $\dexact \in [3.6, 4.8]$
measured in O12--O13 is an \emph{upper bound} on the dynamical
exponent entering the growth equation.
Since $\dexact < 5.0 \ll 7.4$, no aggregation or reweighting
of the existing block-level data can bring the dynamical
exponent into the target window $[7.4, 10.6]$.
\end{corollary}

\begin{proof}
At fixed BFS depth~$n$, the shell size $|S_n|$ is common to all blocks.
Therefore $\Sigma_n^{(c)} = \Delta r_n^{(c)} / |S_n|$ is a monotone
increasing function of $\Delta r_n^{(c)}$ alone.
The weight $w_c = \Delta r_n^{(c)} / \sum_{c'} \Delta r_n^{(c')}$
is likewise a monotone increasing function of the same variable
$\Delta r_n^{(c)}$.
Since both $w_c$ and $\Sigma_n^{(c)}$ are monotone increasing functions
of a single common variable, the Chebyshev sum inequality gives
\[
\sum_c w_c \, \Sigma_n^{(c)}
\;\geq\;
\left(\sum_c w_c\right)
\left(\frac{1}{q-1} \sum_c \Sigma_n^{(c)}\right)
= \Sn,
\]
so $\Reff \geq \Sn$ pointwise in~$n$.
A pointwise lower bound on the observable translates into an upper
bound on the decay exponent: $\alpha_{\mathrm{dyn}} \leq \dexact$.
\end{proof}

This is the sharpest negative result of the paper:
it eliminates the entire class of solutions based on
``better averaging of the blocks.''
If the $\delta$--$\bstar$ tension is to be resolved, the resolution
must come from a modification of the growth equation itself
--- through the shell growth factor, the $q$-normalisation,
or a structural correction $\sigma(q)$ --- not from a refined
treatment of the block statistics.

\section{Re-Derivation of the Growth Equation}
\label{sec:rederivation}

\subsection{Growth equation in the exact-block setting}

In the exact setting, the BFS frontier at step~$n$ contributes new
relational degrees of freedom through all $(q-1)$ Weil blocks
simultaneously.
The total incremental span is
\begin{equation}
\label{eq:totalspan}
\Delta r_n^{\mathrm{tot}}
= \sum_{c} \Delta r_n^{(c)}.
\end{equation}
Each block~$c$ lives in a representation space of dimension~$q$,
so the total available representation space has dimension
$(q-1) \cdot q \sim q^2$.

The Born--Infeld growth bound becomes
\begin{equation}
\label{eq:growthexact}
\frac{\mathrm{d}p}{\mathrm{d}n}
\lesssim c_{\mathrm{BI}} \sqrt{p(n)} \cdot
\frac{\Delta r_n^{\mathrm{tot}}}{(q-1) \cdot q},
\end{equation}
where the denominator $(q-1) \cdot q$ normalises by the total number
of representation degrees of freedom.

\begin{remark}[Status of the $q$-dependent normalisation]
\label{rem:qnorm}
The factor $1/q$ in~\eqref{eq:growthexact} is a
\emph{dimensional ansatz}, not a derived result.
Its motivation is as follows: in the proxy setting, the capacity
is defined in a scalar (one-dimensional) representation, so no
internal dimension enters the normalisation.
In the exact setting, each block has $q$ internal degrees of
freedom, and the fraction of genuinely new directions per
available degree of freedom naturally acquires a factor $1/q$.
This is consistent with the normalisation correction identified
in O14 as the dominant source of the observable-class mismatch.
However, the Born--Infeld constraint $|\partial_t \chi_v| \leq c_{\mathrm{BI}}$
operates at the level of individual relational variables, not at
the representation level.
The correct translation of this per-variable bound into a
per-representation constraint depends on how the representation
degrees of freedom couple to the relational dynamics,
which remains an open derivation.
The factor $1/q$ is the simplest dimensional estimate; a more refined
treatment may yield a different power $q^{-\eta}$ with $\eta \neq 1$.
\end{remark}

\subsection{Separation of variables}

Rewrite~\eqref{eq:growthexact} using
$\Delta r_n^{\mathrm{tot}} / ((q-1) \cdot q)
= \frac{1}{q} \cdot \frac{1}{q-1} \sum_c \Delta r_n^{(c)} / |S_n|
\cdot |S_n|
= \frac{|S_n|}{q} \cdot \Sn$,
where we used $\Sn = \frac{1}{q-1} \sum_c \Sigma_n^{(c)}$
with $\Sigma_n^{(c)} = \Delta r_n^{(c)} / |S_n|$.
Then
\begin{equation}
\label{eq:growthexact2}
\frac{\mathrm{d}p}{\mathrm{d}n}
\lesssim c_{\mathrm{BI}} \sqrt{p(n)} \cdot \frac{|S_n|}{q} \cdot \Sn.
\end{equation}

In the BFS growth regime on the Heisenberg graph,
$|S_n| \sim C_S \, n^{D-1}$ with $D$ the effective growth dimension,
and $\Sn \sim A \, n^{-\dexact}$.
The factor $|S_n| \cdot \Sn \sim C \, n^{D - 1 - \dexact}$.

For the proxy derivation (O6), $|S_n|$ is absorbed into the
definition of~$\Rn$ and does not appear explicitly.
The exact derivation makes the shell growth and the capacity decay
visible as separate factors.

\begin{hypothesis}[Effective growth law under exact-block normalisation]
\label{hyp:growthexact}
Under the ansatz $p(n) \sim A n^\beta$, the exact-block growth
equation~\eqref{eq:growthexact2}, and the dimensional normalisation
ansatz $1/q$ of Remark~\ref{rem:qnorm}, separation of variables gives
\begin{equation}
\label{eq:betaexact}
\beta = \frac{1}{\tfrac{1}{2} + \dexact - (D - 1)
+ \log q / \log n^{*}},
\end{equation}
where $n^{*}$ is the saturation depth.
In the limit $q \to \infty$ with $n^{*} \sim q$, the term
$\log q / \log n^{*} \to 1$.
\end{hypothesis}

\begin{remark}[Status of Hypothesis~\ref{hyp:growthexact}]
\label{rem:status}
This is a heuristic re-derivation, not a theorem.
Two inputs remain conjectural:
(i)~the normalisation factor $1/q$ in~\eqref{eq:growthexact} is a
dimensional ansatz motivated by the representation structure
(each block has $q$ internal degrees of freedom),
but it has not been derived from the Born--Infeld constraint
in the block setting;
(ii)~the term $\log q / \log n^{*}$ is an effective translation of
the $q$-dependent normalisation into exponent space, valid only
at the characteristic depth $n^{*}$.
What is rigorously established is that the proxy formula
$\beta = 1/(1/2 + \delta)$ cannot be applied to $\dexact$ without
modification; the precise form of the modification is the content
of this hypothesis.
\end{remark}

\noindent\textit{Derivation sketch.}
Substituting $p(n) = A n^\beta$ and the power-law forms into
\eqref{eq:growthexact2}:
\[
A \beta n^{\beta - 1}
\lesssim c_{\mathrm{BI}} \sqrt{A} \, n^{\beta/2}
\cdot \frac{C}{q} \, n^{D - 1 - \dexact}.
\]
Collecting powers of~$n$:
$\beta - 1 = \beta/2 + D - 1 - \dexact - \log q / \log n$,
where $\log q / \log n$ accounts for the $q$-dependent normalisation
evaluated at the characteristic depth $n = n^{*}$.
Solving for~$\beta$:
$\beta/2 = \dexact - D + 1 + \log q / \log n^{*}$,
whence $\beta = 1/(\tfrac{1}{2} + \dexact - (D-1) + \log q / \log n^{*})$.
In the proxy limit ($q = 1$, $D = 1$), this reduces to
$\beta = 1/(1/2 + \delta)$, recovering the O6 formula.

\subsection{Two structural scenarios}
\label{sec:scenarios}

\begin{description}[style=nextline]
\item[Scenario C1: Same functional form, corrected exponent]
If the shell growth exponent $D - 1$ and the normalisation term
$\log q / \log n^{*}$ combine to produce an effective correction
that can be absorbed into a redefined exponent
$\deff = \dexact - (D - 1) + \log q / \log n^{*}$,
then~\eqref{eq:betaexact} takes the form
\begin{equation}
\label{eq:C1}
\bstar = \frac{1}{\deff + \tfrac{1}{2}},
\end{equation}
with $\deff$ carrying the full correction.
This is the scenario where the O6 formula survives structurally
but with a different numerical input.

For the Heisenberg graph with $D \approx 4$ (Bass--Guivarc'h
growth exponent), $(D-1) = 3$, and
$\log q / \log n^{*} \approx 1$ at $n^{*} \sim q$,
one obtains $\deff \approx \dexact - 3 + 1 = \dexact - 2$.
With $\dexact \in [3.6, 4.8]$, this gives
$\deff \in [1.6, 2.8]$, yielding $\bstar \in [0.26, 0.48]$.
This is still above the target $(0.09, 0.13)$.

\item[Scenario C2: Modified functional form]
If the corrections cannot be absorbed into a single
redefined exponent, the growth equation acquires an explicit
$q$-dependent structural term:
\begin{equation}
\label{eq:C2}
\bstar = \frac{1}{\deff + \tfrac{1}{2} + \sigma(q)},
\end{equation}
where $\sigma(q)$ encodes the residual effect of the block
structure on the growth dynamics and satisfies
$\sigma(q) \to 0$ as $q \to \infty$.

For C2 to produce $\bstar \in (0.09, 0.13)$ with
$\deff \in [1.6, 2.8]$, the correction must satisfy
$\sigma(q) \in [5.1, 8.6]$, which would require
$\sigma(q) \sim q^{\kappa}$ for some $\kappa > 0$.
Whether such a correction arises from the Weil-block
representation structure is the central open question.
\end{description}

\section{Scenario Discrimination}
\label{sec:discrimination}

\subsection{What would discriminate C1 from C2}

The two scenarios make different predictions for the
$q$-dependence of the dynamical exponent $\alpha_{\mathrm{dyn}}$:

Under C1, $\alpha_{\mathrm{dyn}}$ absorbs all corrections and its
$q$-dependence is fully captured by the shift $\deff = \dexact - 2$.
A measurement of $\Reff$ (Definition~\ref{def:Reff}) at multiple
primes would show power-law decay with an exponent that
\emph{stabilises} as $q$ grows.

Under C2, $\alpha_{\mathrm{dyn}}$ stabilises but $\sigma(q)$
introduces a separate $q$-dependent correction.
The growth equation would show a systematic $q$-dependent
departure from the C1 prediction.

\subsection{Required measurement: block-level redundancy}

The O12/O13 pipeline~\cite{Beau2026a16,Beau2026a17} computes
$\Sigma_n^{(c)}$ per block but discards the complex fingerprint
vectors $z_b^{(c)}(n)$ after Gram--Schmidt orthogonalisation.
The weights $w_c(n)$ of Definition~\ref{def:Reff} are computable
from the existing pipeline outputs ($\Delta r_n^{(c)}$ is stored).

The block-level redundancy measurement therefore requires:
\begin{enumerate}[label=(\roman*)]
\item Compute $w_c(n)$ from the stored $\Delta r_n^{(c)}$ values.
\item Form $\Reff = \sum_c w_c(n) \Sigma_n^{(c)}$.
\item Fit $\Reff \sim B \, n^{-\alpha_{\mathrm{dyn}}}$ and extract
$\alpha_{\mathrm{dyn}}$.
\item Compare $\alpha_{\mathrm{dyn}}$ with $\dexact$ to measure the
aggregation bias.
\item Test whether $\beta = 1/(\alpha_{\mathrm{dyn}} + 1/2)$ falls in
$(0.09, 0.13)$ (C1), or whether an additional $q$-dependent
term is required (C2).
\end{enumerate}

Step~(i) is immediately feasible from existing O12/O13 data.
Steps~(ii)--(v) constitute the minimal computational programme
for discriminating the two scenarios.

\section{Asymptotic Analysis}
\label{sec:asymptotic}

\subsection{The $q \to \infty$ limit}

In the limit $q \to \infty$, three simplifications occur:
\begin{enumerate}[label=(\alph*)]
\item Block universality: $\mathrm{Var}_c(\Sigma_n^{(c)}) \to 0$
by the O13 variance reduction result~\cite{Beau2026a17},
so $\Reff \to \Sn$ and $\alpha_{\mathrm{dyn}} \to \dexact$.
\item The normalisation correction $\log q / \log n^{*} \to 1$.
\item The correction $\epsilon(q) \to 0$ in~\eqref{eq:O14corrected}.
\end{enumerate}
In this limit, the growth equation reduces to
$\beta = 1/(1/2 + \dexact - 2)$ under C1,
or $\beta = 1/(1/2 + \dexact - 2 + \sigma(\infty))$ under C2.

If $\sigma(\infty) = 0$, C2 reduces to C1 asymptotically.
If $\sigma(\infty) > 0$, C2 introduces a permanent structural
correction even in the continuum limit.

\subsection{Finite-$q$ phenomenology}

At accessible primes ($q \leq 211$), the correction terms are
non-negligible.
Table~\ref{tab:scenarios} reports the predicted $\bstar$ under
both scenarios for the five O12/O13 primes.

\begin{table}[ht]
\centering
\caption{Predicted $\bstar$ under scenarios C1 and C2 for the
five O12/O13 primes.
C1 uses $\deff = \dexact - 2$;
C2 requires $\sigma(q)$ to reach the target.}
\label{tab:scenarios}
\begin{tabular}{ccccc}
\toprule
$q$ & $\dexact$ & $\deff$ (C1) & $\bstar$ (C1) & $\sigma(q)$ needed (C2) \\
\midrule
29 & 4.37 & 2.37 & 0.348 & 5.4 \\
61 & 4.80 & 2.80 & 0.303 & 5.9 \\
101 & 4.28 & 2.28 & 0.360 & 5.3 \\
151 & 3.90 & 1.90 & 0.417 & 5.0 \\
211 & 3.59 & 1.59 & 0.478 & 4.7 \\
\bottomrule
\end{tabular}
\end{table}

Under C1, $\bstar \in [0.30, 0.48]$ for all tested primes:
closer to the target than the uncorrected value
$\bstar \in [0.19, 0.21]$ from O12, but still above $(0.09, 0.13)$
by a factor of approximately~3.

Under C2, the required correction $\sigma(q) \in [4.7, 5.9]$
is large and shows a mild $q$-dependence.
Whether a correction of this magnitude can arise from the
Weil-block structure is an open question.

\section{Open Directions}
\label{sec:open}

\begin{description}[style=nextline]
\item[O15-O1 (Block-level redundancy extraction)]
Compute $\Reff$ from the O12/O13 pipeline outputs and extract
$\alpha_{\mathrm{dyn}}$ at $q \in \{29, 61, 101, 151, 211\}$.
This is the minimal measurement required to discriminate C1 from C2.

\item[O15-O2 (Shell growth exponent)]
The effective growth dimension $D \approx 4$ (Bass--Guivarc'h)
is an asymptotic estimate.
The exact value of $D$ in the pre-saturation window
$[n_0, n_1]$ at finite~$q$ may differ.
Extracting $D$ directly from the BFS data would sharpen the
C1 prediction.

\item[O15-O3 (Representation-theoretic $\sigma(q)$)]
Under C2, the correction $\sigma(q)$ must be derived from the
representation structure of the Heisenberg group.
The Weil representation has a metaplectic phase structure
($\mathrm{SL}_2(\mathbb{Z}/q\mathbb{Z})$ acts via Weil's
representation) that may contribute a systematic correction
to the growth dynamics.

\item[O15-O4 (Quark and neutrino sectors)]
Both scenarios modify the $\delta \to \bstar$ map.
The extension to the quark and neutrino sectors
(O14-O2~\cite{Beau2026a18}) requires re-evaluating the
corrected relation at the block structures appropriate
for those sectors.
\end{description}

\section{Conclusion}
\label{sec:conclusion}

The O3--O7 derivation chain that produces
$\bstar = 1/(\delta + 1/2)$ is logically correct in its domain:
a scalar global capacity on proxy-level LPS dynamics.
The failure is not an error but a scope limitation:
the scalar-observable hypothesis H-O6-1 does not hold for the
exact Weil-block capacity measured in O12--O14.

The non-equivalence theorem (Proposition~\ref{prop:noneq})
establishes that the block-averaged exponent $\dexact$ does not
coincide with the scalar exponent~$\alpha$ entering the growth
equation, and that the bias goes in the wrong direction
(Remark~\ref{rem:bias}).

The effective dynamical redundancy $\Reff$
(Definition~\ref{def:Reff}) is the correct observable entering
the growth equation in the exact setting.
Its exponent $\alpha_{\mathrm{dyn}}$ is smaller than $\dexact$,
so the aggregation correction alone cannot resolve the tension.

The aggregation no-go (Corollary~\ref{cor:nogo}) is the sharpest
negative result: no reweighting of the block-level data can bring
the dynamical exponent into the target window.
The resolution must come from the growth equation itself.

The re-derivation of the growth equation
(Hypothesis~\ref{hyp:growthexact}) reveals that the passage from
proxy to exact introduces two additional corrections:
the shell growth dimension $(D - 1)$ and the block-normalisation
$\log q / \log n^{*}$.
These corrections reduce the effective exponent and bring
$\bstar$ closer to the target, but under Scenario~C1 the residual
gap is still a factor of~3.

Scenario~C2, which introduces a $q$-dependent structural correction
$\sigma(q)$, can close the gap but requires $\sigma(q) \sim 5$,
a large value whose derivation from the Weil representation
structure is an open problem.

The minimal next step is the block-level redundancy extraction
(O15-O1): computing $\Reff$ from existing O12/O13 data to measure
$\alpha_{\mathrm{dyn}}$ directly and discriminate the two scenarios.
This measurement is feasible with the current pipeline outputs and
constitutes the critical path forward.

The O-series has now traced the chain from obstruction identification
(O8--O9) through exact measurement (O10--O13), theoretical
correspondence (O14), and derivation audit (O15).
The status of Source~S2 is transformed:
it is no longer an unexplained numerical tension,
nor a theoretical impasse,
but a precisely formulated computational problem.
The gap between $\dexact$ and the target window is quantified,
its origin is localised at the growth-equation level,
and the discriminating measurement (O15-O1: extraction of
$\alpha_{\mathrm{dyn}}$ from existing pipeline data) is
immediately feasible.
O15 thus converts the open question ``why does $\delta$ not
match $\bstar$?'' into the concrete programme
``measure $\Reff$, test C1, and if C1 fails,
derive $\sigma(q)$ from the Weil representation structure.''

\bibliographystyle{plain}
      \bibliography{cosmochrony-bibliography}

\end{document}
